### 2018-05-08 Paper/examples.tex
..

In section~\ref{sec:method}, we presented a methodology for screening model
parameters using an approximate estimate of the upper bound on the Sobol'
total effect sensitivity index. The screening process helped identify parameters
that are relatively insignificant with respect to the observable. These insignificant
parameters could thus be fixed at their nominal values thereby reducing the
dimensionality of the problem.

In this section, we motivate this approach based
on parameter screening to construct low-dimensional PC surrogates by 
applying it to simple test problems: (1) Borehole function, (2) Non-linear Oscillator, and
(3) Semi-linear Elliptic PDE. Since model evaluations in all these test
problems are not expensive, it allows us to compare relative importance of
model parameters,
obtained using the derivative-based sensitivity approach with that obtained
using converged estimates of Sobol' total effect sensitivity index, $\mathcal{T}(\theta_i)$.

%\cite{Sobol:2001}:
%\be
%\mathcal{T}(\theta_i) = 
%\frac{\mathbb{E}_{\bm{\theta}\sim i}[\mathbb{V}_{\theta_i}(\mathcal{G}|\bm{\theta}_{\sim i})]}{\mathbb{V}(\mathcal{G})}
%\ee

\noindent $\mathcal{T}(\theta_i)$ is essentially the ratio of the contribution to the variance of an
observable $\mathcal{G}$ by the parameter $\theta_i$ to the total variance of $\mathcal{G}$.
Additionally, we compare convergence of the PCE constructed in
the full parameter space with those constructed in the reduced space. To assess convergence, 
we rely on the 
leave-one-out-cross-validation error ($\epsilon_{\mbox{\tiny LOO}}$), defined
as follows~\cite{Blatman:2010}:

\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum_{i=1}^N\left(\mathcal{Y}^{\mbox{\tiny M}}(\bm{\theta}_i) - 
 \mathcal{Y}^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum_{i=1}^N
\left(\mathcal{Y}^{\mbox{\tiny M}}(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee

\noindent where $\widetilde{\mu}~=~\frac{1}{N}\sum_{i=1}^N \mathcal{Y}^{\mbox{\tiny M}}(\bm{\theta}_i)$
is the sample mean of the model response, and $ \mathcal{Y}^{\mbox{\tiny {PC}\textbackslash i}}$
is the PCE surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$ model realization. 
From Eq.~\ref{eq:loo}, it appears that $N$ PCE's are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.
However, in practice a modified formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009},
independent of $\mathcal{Y}^{\mbox{\tiny {PC}\textbackslash i}}$ is used.

The converged low-dimensional PCE is assessed
for predictive accuracy. While several techniques could be used, we consider predictive verification
in an L-2 sense as well as a probabilistic sense. In the former case, we estimate a relative L-2 norm
of the error ($\epsilon_{\mbox{\tiny{L-2}}}$) as follows:

\be
\epsilon_{\mbox{\tiny{L-2}}} = 
\frac{\left[\sum_{i=1}^N\left(\mathcal{Y}^{\mbox{\tiny M}}(\bm{\theta}_i) - 
 \mathcal{Y}^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum_{i=1}^N
\left(\mathcal{Y}^{\mbox{\tiny M}}(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee

\noindent As discussed earlier in section~\ref{sec:method}, computational cost associated with
L-2 verification is negligible since function evaluations or model realizations obtained during the
screening step could be re-used to estimate $\epsilon_{\mbox{\tiny{L-2}}}$ using Eq.~\ref{eq:l2}.
For the purpose of probabilistic verification, we compare the probability density function of
the observable obtained using PCE's in the full space and the reduced space in the following
test problems. In situations where model runs are computationally expensive, probabilistic verification might
not be feasible.

>>>>

In section~\ref{sec:method}, we presented a methodology for constructing
an ROS by identifying unimportant parameters based on 
estimates of the upper bound on the Sobol'
total effect sensitivity index ($\mathcal{T}(\theta_i)$) for individual parameters.
In this section, we motivate the methodology by applying it to three test problems,
namely, the borehole function, a non-linear oscillator, and a semi-linear elliptic PDE.
Since model evaluations in all these test
problems are inexpensive, we compare relative importance of model parameters based
on DGSMs, obtained using model predictions with 
converged estimates of $\mathcal{T}(\theta_i)$ based on the surrogate, constructed in the
full parameter space. Additionally, to illustrate computational gains, we compare
convergence trends as a function of training runs for the ROS with the full-space
surrogate. Furthermore, as discussed earlier in section~\ref{sec:method}, we compare
PDFs of the model output, obtained using the ROS and the full-space surrogate for
the purpose of verification.

### 2018-05-08 Paper/intro.tex
..

The emerging field of uncertainty quantification (UQ) aims at methodologies for 
incorporating, characterizing, quantifying, propagating, and reducing the 
uncertainties associated with predictive models and simulations. For situations
involving complex physical models and compute-intensive simulations, an
efficient approach to construction of model surrogates is sought to enable UQ
in a tractable manner. Polynomial chaos expansion 
(PCE)~\cite{Xiu:2002,Ghanem:2003,Eldred:2008,Olivier:2010} and 
Gaussian Process (GP)~\cite{Rasmussen:2004} or Kriging~\cite{Stein:2012} are
among the most commonly used surrogates for scientific applications. However,
both approaches quickly become prohibitive when the set of uncertain model
inputs and parameters is large-dimensional. Specifically, in the case of a PCE,
a $d$-dimensional polynomial basis with a total order of truncation, $p$ 
requires model realizations at $(p+1)^d$ quadrature nodes to exactly determine the PC
coefficients using a fully-tensorized Gauss quadrature. Consider a case where $d$ = 8 which
could be a regarded as a low-dimensional problem, and $p$ = 5, requires model evaluations at
roughly 1.7 million quadrature nodes. However, the degree of exactness in numerical estimation 
of the coefficients as made possible with a fully-tensored quadrature might not be desired. In that
case, computational effort can be reduced considerably by using nested
quadratures~\cite{Gentleman:1972a,Gentleman:1972b,Novak:1999,Waldvogel:2006} which allow 
model realizations at a given level to be used at higher levels of accuracy as well as sparse grids
based on Smolyak tensorization~\cite{Smolyak:1963}. Additonally, sparse basis 
techniques~\cite{Peng:2014,Hampton:2015,Blatman:2011} have been used to reduce  the required
computational effort for constructing a reasonably accurate PC surrogate.

One of the primary objectives for constructing a model surrogate is to be able to perform 
parametric sensitivity analysis tractably. Sensitivity analysis provides an insight into the
relative contributions of the uncertain model parameters and inputs to the uncertainty in
predictions. The analysis could potentially reduce the dimensionality of the problem.
However, as discussed earlier, constructing the surrogate might still require a large amount
of computational resources. It can thus be understood that dimension reduction prior to
constructing the surrogate might lead to enormous computational savings. In this work,
we present a strategy for identifying and screening uncertain model parameters that 
are significantly less important that the rest, thereby reducing the dimensionality of the
problem and enabling the construction of a reduced-order surrogate. As discussed in
section~\ref{sec:method}, parameter ranks are determined using a \textit{screening-metric}
based on derivative-based global sensitivity 
measures~\cite{Sobol:2009,Sobol:2010,Lamboni:2013,Kucherenko:2009,Kucherenko:2016}. 
The strategy is agnostic to the choice of methodology for constructing the surrogate. However, 
we will be using PCE to demonstrate the suitability of the proposed strategy. Below, in 
section~\ref{sec:bg}, we provide a brief introduction to derivative-based global sensitivity analysis
in the present context as well as the polynomial chaos methodology. Later, in section~\ref{sec:examples},
we motivate the proposed strategy by implementing it to a suite of simple applications, namely, the 
classic borehole function, a non-linear oscillator, and a semi-linear elliptic PDE. Finally, we demonstrate
suitability of our approach by using it to reduce the dimensionality of a H$_2$/O$_2$ kinetics problem
involving 19 reactions in section~\ref{sec:app}.

>>>>

The emerging field of uncertainty quantification (UQ) aims at methodologies for 
incorporating, characterizing, quantifying, propagating, and reducing the 
uncertainties associated with predictive models and simulations. For situations
involving complex physical models and compute-intensive simulations, an
efficient approach to construction of model surrogates is sought to enable UQ
in a tractable manner. Polynomial chaos expansion 
(PCE)~\cite{Xiu:2002,Ghanem:2003,Olivier:2010} and 
Gaussian Process (GP)~\cite{Rasmussen:2004} or Kriging~\cite{Stein:2012} are
among the most commonly used surrogates for scientific applications. However,
both approaches quickly become prohibitive when the set of uncertain model
inputs and parameters is large-dimensional. Specifically, in the case of a PCE,
a $d$-dimensional polynomial basis with a total order of truncation, $p$ 
requires model realizations at $(p+1)^d$ quadrature nodes to exactly determine the PC
coefficients using a fully-tensorized Gauss quadrature. Consider a case where $d$ = 8 which
could be a regarded as a low-dimensional problem, and $p$ = 5, requires model evaluations at
roughly 1.7 million quadrature nodes. However, the degree of exactness in numerical estimation 
of the coefficients as made possible with a fully-tensored quadrature might not be desired. In that
case, computational effort can be reduced considerably by using nested
quadratures~\cite{Gentleman:1972a,Gentleman:1972b,Novak:1999,Waldvogel:2006} which allow 
model realizations at a given level to be used at higher levels of accuracy as well as sparse grids
based on Smolyak tensorization~\cite{Smolyak:1963}. Additonally, sparse basis 
techniques~\cite{Peng:2014,Hampton:2015,Blatman:2011} have been used to reduce  the required
computational effort for constructing a reasonably accurate PC surrogate.

One of the primary objectives for constructing a model surrogate is to be able to perform 
parametric sensitivity analysis tractably. Sensitivity analysis provides an insight into the
relative contributions of the uncertain model parameters and inputs to the uncertainty in
predictions. The analysis could potentially reduce the dimensionality of the problem.
However, as discussed earlier, constructing the surrogate might still require a large amount
of computational resources. It can thus be understood that dimension reduction prior to
constructing the surrogate might lead to enormous computational savings. In this work,
we present a strategy for identifying and screening uncertain model parameters that 
are significantly less important that the rest, thereby reducing the dimensionality of the
problem and enabling the construction of a reduced-order surrogate (ROS). As discussed in
section~\ref{sec:method}, parameter ranks are determined using a \textit{screening-metric}
based on derivative-based global sensitivity 
measures~\cite{Sobol:2009,Sobol:2010,Lamboni:2013,Kucherenko:2009,Kucherenko:2016}. 
The strategy is agnostic to the choice of methodology for constructing the surrogate. However, 
we will be using PCE to demonstrate the suitability of the proposed strategy. Below, in 
section~\ref{sec:bg}, we provide a brief introduction to derivative-based global sensitivity analysis
in the present context as well as the polynomial chaos methodology. Later, in section~\ref{sec:examples},
we motivate the proposed strategy by implementing it to a suite of simple applications, namely, the 
classic borehole function, a non-linear oscillator, and a semi-linear elliptic PDE. Finally, we demonstrate
suitability of our approach by using it to reduce the dimensionality of a H$_2$/O$_2$ kinetics problem
involving 19 reactions in section~\ref{sec:app}.

### 2018-05-08 Paper/method.tex
..

In this section, we outline the underlying methodology for constructing a 
reduced-order surrogate using sensitivity analysis. The term `reduced-order'
in the present context implies that the surrogate is constructed in a 
subspace that sufficiently captures the uncertainty in the model output. 
For instance, in the case of PC surrogates, the dimensionality of the polynomial
basis functions is effectively reduced. A possible approach to 
constructing a reduced-order surrogate involves sensitivity analysis of the
uncertain model parameters with respect to a given output, and thereby
disregarding the uncertainty associated with parameters considered as unimportant.
For this purpose, a global sensitivity analysis based on Sobol indices could be pursued.
However, as discussed earlier in sections~\ref{sec:intro} and~\ref{sec:bg}, determining converged 
estimates of Sobol sensitivity indices typically requires tens of thousands
of model evaluations. Consequently, the exercise becomes prohibitive in
situations where the model runs are compute-intensive. Instead, we estimate a normalized
upper-bound ($\widehat{\mathcal{C}_i\mu_i}$, see Eq.~\ref{eq:bound}) on the Sobol total
effect index ($\mathcal{T}(\theta_i)$) for each parameter, $\theta_i$, and use it as a 
metric to identify the unimportant parameters. Below, we provide an algorithm for 
parameter screening. It is assumed that the partial derivatives in Eq.~\ref{eq:mu}
are estimated using finite difference.

@@

\begin{breakablealgorithm}
  \caption{Parameter screening with derivative-based sensitivity measures.}
  \begin{algorithmic}[1]
    \Procedure{Screening}{}
      \State Generate $n_1$ points in $\mathbb{R}^{d}$.\Comment{$d$: 
             Number of uncertain model parameters.}
      \State Perturb each point along the $d$ directions to obtain a set of $n_1(d+1)$ points.
      \State Compute $\mu_i$ using model evaluations at the $n_1(d+1)$ points in Eq.~\ref{eq:mu}
      \State Determine initial ranks, $\mathcal{R}^{old}$ based on $\widehat{\mathcal{C}_i\mu_i}$ values for $\theta_i$.
      \State set $k$ = 1\Comment{Iteration counter}
      \Do
        \State Generate $n_k$ new points in $\mathbb{R}^{d}$.
        \State Perturb each point along the $d$ directions to obtain a set of $n_k(d+1)$ points.
        \State Compute and store model evaluations at the $n_k(d+1)$ points.
        \State Compute $\mu_i$ using prior model evaluations at $(d+1)(n_1 + \sum_j^k n_j)$ points.
        \State Determine new ranks, $\mathcal{R}^{new}$ based on updated $\widehat{\mathcal{C}_i\mu_i}$ values.
        \State Compute $max\_pdev$ = max$\left(\frac{|\mu_{i,k} - 
               \mu_{i,k-1}|}{ \mu_{i,k-1}}\right)$.\Comment{$max\_pdev$:
               Maximum percentage deviation in $\mu_i$ between successive iterations.}
        \State set $k$ = $k$ + 1
      \doWhile{($\mathcal{R}^{\tiny{new}}$ $\neq$ $\mathcal{R}^{\tiny{old}}$ {\bf or}  
               $max\_pdev$~$>~\tau$)\Comment{$\tau$:~Tolerance}}
    \EndProcedure
  \end{algorithmic}
\end{breakablealgorithm}

@@

Based on the steps outlined in the above algorithm, it can be said that the set of sample points used for
estimating the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks
between successive iterations as well as a certain degree of convergence in estimates of the screening 
metric for each uncertain parameter is accomplished. Parameter
screening is an integral part of the overall strategy for constructing the reduced-order surrogate as 
depicted below using a flow-diagram.

@@

\begin{center}
\begin{tikzpicture}[node distance=1.5cm]

@@

\draw [line] (start) -- (qoi);

\node (tol) [process, below of=qoi, text width=8em] {Set an initial tolerance, $\tau$};

\draw [line] (qoi) -- (tol);

@@

\draw [line] (tol) -- (screen);

\node (dr) [draw, diamond, aspect=2.0] {Is \\ Dimension-reduction \\ possible$?$};

@@

>>>>

In this section, we outline the underlying methodology for constructing a 
reduced-order surrogate (ROS) using sensitivity analysis. The term `reduced-order'
in the present context implies that the surrogate is constructed in a 
subspace that sufficiently captures the uncertainty in the model output. 
In the case of PC surrogates used in this work, it implies that the dimensionality
of the polynomial basis functions is effectively reduced.

A possible approach to 
constructing the ROS involves sensitivity analysis of the
uncertain model parameters with respect to a given output, and thereby
disregarding the uncertainty associated with parameters considered as unimportant.
For this purpose, a global sensitivity analysis based on Sobol indices could be pursued.
However, as discussed earlier in sections~\ref{sec:intro} and~\ref{sec:bg}, determining converged 
estimates of Sobol sensitivity indices typically requires tens of thousands
of model evaluations. Consequently, the exercise becomes prohibitive in
situations where the model runs are compute-intensive. Instead, we adopt a novel approach
that focuses on estimating the upper-bound
upper-bound ($\widehat{\mathcal{C}_i\mu_i}$, see Eq.~\ref{eq:bound}) on Sobol total
effect index ($\mathcal{T}(\theta_i)$) for each parameter, $\theta_i$, and using it as a 
metric to screen parameters that are relatively unimportant. Below, we provide a generalized
strategy for parameter screening.

@@

\begin{breakablealgorithm}
  \caption{Parameter screening with DGSMs: A generalized approach.}
  \begin{algorithmic}[1]
    \Procedure{Screening}{}
      \State Generate $n_1$ samples in $\mathbb{R}^{N_p}$ according to the joint probability density, $\bm{f(\theta)}$.
      \State Evaluate the gradient vector, $\nabla_{\bm{\theta}}G_k$ at each sample point, $\bm{\theta}_k$.
      \State Approximate sensitivity measures, $\bm{\mu}$: \be\bm{\mu} \approx \hat{\bm{\mu}} = 
      \frac{1}{n_1}\sum\limits_{k=1}^{n_1}(\nabla_{\bm{\theta}}G_k)(\nabla_{\bm{\theta}}G_k)^{T}\nonumber\ee
      \State Estimate $\widehat{\mathcal{C}_i\mu_i}$ for each parameter, $\theta_i$ and determine initial ranks,
       $\mathcal{R}^{old}$.
      \State set $s$ = 1\Comment{Iteration counter}
      \Do
        \State Generate $n_s$ new points in $\mathbb{R}^{N_p}$.
        \State Determine new ranks, $\mathcal{R}^{new}$ based on updated $\widehat{\mathcal{C}_i\mu_i}$ values.
        \State Compute $\Delta\mu_m$ = max$\left(\frac{|\mu_{i,s} - 
               \mu_{i,s-1}|}{ \mu_{i,s-1}}\right)$.\Comment{$\Delta\mu_m$:
               Maximum deviation in $\mu_i$ between successive iterations.}
        \State set $s$ = $s$ + 1
      \doWhile{($\mathcal{R}^{\tiny{new}}$ $\neq$ $\mathcal{R}^{\tiny{old}}$ {\bf or}  
               $\Delta\mu_m$~$>~\tau$)\Comment{$\tau$:~Tolerance}}
    \EndProcedure
  \end{algorithmic}
\end{breakablealgorithm}

@@

In case the gradient, $\nabla_{\bm{\theta}}G_k$ cannot be evaluated analytically, we could use finite-difference
to approximate its value as shown in Eq.~\ref{eq:partial}. In that case, the above algorithm could be adapted as
follows:

@@

\begin{breakablealgorithm}
  \caption{Parameter screening with DGSMs using finite differences.}
  \begin{algorithmic}[1]
    \Procedure{Screening}{}
      \State Generate $n_1$ samples in $\mathbb{R}^{N_p}$ according to the joint probability density, $\bm{f(\theta)}$.
      \State Perturb each point along the $N_p$ directions to obtain a set of $n_1(N_p+1)$ samples.
      \State Estimate $\nabla_{\bm{\theta}}G_k$ using model evaluations at the $n_1$ samples and the corresponding 
                 neighbouring samples for each component of $\bm{\theta}$.
      \State Approximate sensitivity measures, $\bm{\mu}$: \be\bm{\mu} \approx \hat{\bm{\mu}} = 
      \frac{1}{n_1}\sum\limits_{k=1}^{n_1}(\nabla_{\bm{\theta}}G_k)(\nabla_{\bm{\theta}}G_k)^{T}\nonumber\ee
      \State Estimate $\widehat{\mathcal{C}_i\mu_i}$ for each parameter, $\theta_i$ and determine initial ranks,
       $\mathcal{R}^{old}$.
      \State set $s$ = 1\Comment{Iteration counter}
      \Do
        \State Generate $n_s$ new points in $\mathbb{R}^{N_p}$.
        \State Perturb each point along the $N_p$ directions to obtain a set of $n_s(N_p+1)$ points.
        \State Compute $\mu_i$ using model evaluations at $(N_p+1)(n_1 + \sum\limits_j^s n_j)$ points.
        \State Determine new ranks, $\mathcal{R}^{new}$ based on updated $\widehat{\mathcal{C}_i\mu_i}$ values.
        \State Compute $\Delta\mu_m$ = max$\left(\frac{|\mu_{i,s} - 
               \mu_{i,s-1}|}{ \mu_{i,s-1}}\right)$.\Comment{$\Delta\mu_m$:
               Maximum deviation in $\mu_i$ between successive iterations.}
        \State set $s$ = $s$ + 1
      \doWhile{($\mathcal{R}^{\tiny{new}}$ $\neq$ $\mathcal{R}^{\tiny{old}}$ {\bf or}  
               $\Delta\mu_m$~$>~\tau$)\Comment{$\tau$:~Tolerance}}
    \EndProcedure
  \end{algorithmic}
\end{breakablealgorithm}

%\bigskip
%
%\begin{breakablealgorithm}
%  \caption{Parameter screening with derivative-based sensitivity measures.}
%  \begin{algorithmic}[1]
%    \Procedure{Screening}{}
%      \State Generate $n_1$ points in $\mathbb{R}^{d}$.\Comment{$d$: 
%             Number of uncertain model parameters.}
%      \State Perturb each point along the $d$ directions to obtain a set of $n_1(d+1)$ points.
%      \State Compute $\mu_i$ using model evaluations at the $n_1(d+1)$ points in Eq.~\ref{eq:mu}
%      \State Determine initial ranks, $\mathcal{R}^{old}$ based on $\widehat{\mathcal{C}_i\mu_i}$ values for $\theta_i$.
%      \State set $k$ = 1\Comment{Iteration counter}
%      \Do
%        \State Generate $n_k$ new points in $\mathbb{R}^{d}$.
%        \State Perturb each point along the $d$ directions to obtain a set of $n_k(d+1)$ points.
%        \State Compute and store model evaluations at the $n_k(d+1)$ points.
%        \State Compute $\mu_i$ using prior model evaluations at $(d+1)(n_1 + \sum_j^k n_j)$ points.
%        \State Determine new ranks, $\mathcal{R}^{new}$ based on updated $\widehat{\mathcal{C}_i\mu_i}$ values.
%        \State Compute $max\_pdev$ = max$\left(\frac{|\mu_{i,k} - 
%               \mu_{i,k-1}|}{ \mu_{i,k-1}}\right)$.\Comment{$max\_pdev$:
%               Maximum percentage deviation in $\mu_i$ between successive iterations.}
%        \State set $k$ = $k$ + 1
%      \doWhile{($\mathcal{R}^{\tiny{new}}$ $\neq$ $\mathcal{R}^{\tiny{old}}$ {\bf or}  
%               $max\_pdev$~$>~\tau$)\Comment{$\tau$:~Tolerance}}
%    \EndProcedure
%  \end{algorithmic}
%\end{breakablealgorithm}
%
%\bigskip

Based on the steps outlined in the above algorithm, it can be said that the set of sample points used for
estimating the screening metric, $\widehat{\mathcal{C}_i\mu_i}$ is enriched until consistency in ranks
between successive iterations as well as a certain degree of convergence in estimates of the screening 
metric for each uncertain parameter is accomplished. Parameter
screening is an integral part of the overall strategy for constructing the ROS as 
depicted below using a flow-diagram.

\begin{figure}[htbp]
\begin{center}

\tikzset{
    arro/.style={{Square[]->}}
}

\begin{tikzpicture}[node distance=1.5cm]

@@

\draw [arro] (start) -- (qoi);

\node (tol) [io, below of=qoi, text width=7em,align=left,yshift=-0.2cm] {Set an initial tolerance, $\tau$};

\draw [arro] (qoi) -- (tol);

@@

\draw [arro] (tol) -- (screen);

\node (dr) [draw, diamond, aspect=1.5,yshift=-6.9cm, text width=4.5em, text centered] {Is DR possible$?$};

\draw [arro] (screen) -- (dr);

\node (fss) [process, right of=dr, text width=9em,xshift=3.5cm] {Construct full-space surrogate (FSS)};

\draw [arro] (dr) -- node[anchor=south] {N} (fss);

\node (res) [draw, diamond, aspect=1.5, xshift=3.5cm,text width=4.5em, text centered, right of=fss] {FSS converged$?$};

\draw [arro] (fss) -- (res);

\node (dtol) [io, above of=res, text width=6em,align=left,yshift=1.2cm,text centered]
 {\vspace{-5mm}\be\tau = \tau - \delta\tau,~\delta\tau > 0\nonumber\ee};

\draw [arro] (dtol) -| (screen);

 \draw [arro] (res) -- node[anchor=west,yshift=-0.1cm] {N} (dtol);

 \node (assess) [process, below of=dr, text width=11em,yshift=-1.2cm] {If applicable, assess $\alpha$ and convergence
  of FSS};

 \draw [arro] (dr) -- node[anchor=east] {Y}(assess);

 \node (con) [draw, diamond, aspect=1.5,text width=4em, text centered, below of=assess,yshift=-0.7cm] {Continue$?$};

 \draw [arro] (assess) -- (con);

\node (ros) [process, below of=con, text width=6em, yshift=-0.7cm] {Construct the ROS};

 \draw [arro] (con) -- node[anchor=east] {Y}(ros);

\node (verify) [process, below of=ros, text width=9em,yshift=-0.2cm] {Verify accuracy of the surrogate};

\draw [arro] (ros) -- (verify);

\node (stop) [startstop, right of=verify,xshift=3.5cm] {Stop};

\draw [arro] (verify) -- (stop);

\draw [arro] (res) |- node[anchor=east,yshift=6.9cm] {Y} (stop);

\draw [arro] (con) -| node[anchor=south,xshift=-3cm] {N} (stop);

@@

\caption{Flow-diagram outlining the overall strategy for constructing reduced-order surrogates using
DGSMs. Note that DR is an abbreviation for dimension reduction.}
\label{fig:flow}
\end{figure}

As illustrated in the above flow-diagram, application of the above methodology to construct an ROS 
relies on the choice of an \textit{appropriate} model output. Since the parameters are screened
using DGSMs, the model output should exhibit a smooth dependence on the
individual uncertain parameters to enable a reasonable approximation of the derivatives in Eq.~\ref{eq:mu}.
Thus, for a given problem involving multiple outputs or model predictions, it is likely that a reduced-order 
surrogate might be possible only for selected outputs. Hence, where possible, it is recommended to a priori assess the
nature of dependence of a given model output on the parameters.

The amount of computational effort associated with the presented methodology can be mainly attributed to
two steps: I. Parameter Screening, and II. Constructing a converged ROS. Computational gains are realized
in situations where constructing the surrogate in the full parameter space is more expensive than the combined cost
associated with these steps. Determining the optimal allocation of computational resources for these steps, however, 
is not possible a priori. Hence, we suggest estimating the DGSMs on a small set of samples (say, $n_1$ = 5) 
initially, and screening with a relatively large value of tolerance, $\tau$ while keeping a track of computational expense.
Tighter tolerances on $\bm{\mu}$ are
gradually imposed until the screening process yields conclusive results pertaining to the possibility of dimension-reduction
or a reasonably converged surrogate is constructed using the available set of model evaluations in the full parameter space. 
Hence, the combined computational cost associated with steps I and II could be less, equal to, or greater than that needed
to construct the surrogate in the full parameter space. To avoid the case of greater computational effort, we suggest 
that the decision to construct the ROS should account for the degree of dimension-reduction predicted at the end of the
screening step ($\alpha$, defined below in Eq.~\ref{eq:alpha}), and the state of convergence of the full-space surrogate. 
Enormous computational savings are in fact possible when dealing with complex, 
compute-intensive models, even if $\alpha$ in the following equation is
not too large.

\be
\alpha = \frac{\mbox{Dimensionality of ROS}}{\mbox{Dimensionality of the full surrogate}}
\label{eq:alpha}
\ee

In order to assess convergence of the ROS, we estimate the leave-one-out cross validation error as
follows:

\be
\epsilon_{\mbox{\tiny LOO}} = 
\frac{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - G^{\mbox{\tiny {PC}\textbackslash i}}(\bm{\xi(\theta}_i))\right)^2}
{\sum\limits_{i=1}^{N_l}\left(G(\bm{\theta}_i) - \widetilde{\mu}\right)^2}
\label{eq:loo}
\ee

\noindent where $N_l$ is the number of training points, 
$\widetilde{\mu}~=~\frac{1}{N}\sum\limits_{i=1}^N G(\bm{\theta}_i)$
is the sample mean of the model response, and $ G^{\mbox{\tiny {PC}\textbackslash i}}$
is the PCE surrogate constructed using all but the $i^{\mbox{\tiny{th}}}$ model realization. 
From Eq.~\ref{eq:loo}, it appears that $N$ PCEs are needed to evaluate $\epsilon_{\mbox{\tiny LOO}}$.
However, in practice a modified formulation for $\epsilon_{\mbox{\tiny LOO}}$~\cite{Blatman:2009},
independent of $\mathcal{Y}^{\mbox{\tiny {PC}\textbackslash i}}$ is used.

Finally, as outlined in Figure~\ref{fig:flow}, the converged ROS is verified for predictive accuracy. 
To this end, we compute a relative L-2 norm of the difference in predictions between the model
and the ROS ($\epsilon_{\mbox{\tiny{L-2}}}$) as follows:

\be
\epsilon_{\mbox{\tiny{L-2}}} = \frac{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i) - 
G^{\mbox{\tiny {PC}}}(\bm{\xi(\theta}_i))\right)^2\right]^{\frac{1}{2}}}
{\left[\sum\limits_{i=1}^{N_v}\left(G(\bm{\theta}_i)\right)^2\right]^{\frac{1}{2}}}
\label{eq:l2}
\ee

\noindent where $N_v$ is the number of sampling points in the full parameter space at which model
evaluations are available. It must be noted that the computational effort associated with estimating
 $\epsilon_{\mbox{\tiny{L-2}}}$
is negligible since we exploit the available set of model evaluations in the full parameter space, generated
during the screening step. Additionally, probability density function (PDF) based on these full-space 
model evaluations are compared with the PDF generated using a large set of evaluations of the same
output from the ROS.
